---
title: "Untitled"
output: ''
---

Load different packages that will use in this model:

```{r}
pkg_list = c("ggplot2", "plyr", "stringr")
mia_pkgs = pkg_list[!(pkg_list %in% installed.packages()[,"Package"])]
if(length(mia_pkgs) > 0) install.packages(mia_pkgs)
loaded_pkgs = lapply(pkg_list, require, character.only=TRUE)
```

Set up the data:
```{r}
jdata = read.csv("datarelatedjobs.csv")
names(jdata)
head(jdata)
```

Overiew about our dataset:

```{r}
summary(jdata)
```

```{r}
count(jdata$Standing)
```

1. Comparing between companies in the North and the South of California (using side-by-side histogram):

- Number of companies in North vs South
- Average minimum experience. 
- Number of Bachelor's, Master's, Ph.D that companies recruiting in North vs South. (All levels = add 1 to every value)
- Average salary in North vs South

Set up 2 subset for North and South:

```{r}
North_sub = jdata[which(jdata$Region == "North CA"),]
head(North_sub,20)
South_sub = jdata[which(jdata$Region == "South CA"),]
head(South_sub,20)

```

Number of companies:

```{r}
count(North_sub$Company)$freq
n_companies_north = rep(0, length(count(North_sub$Company)$freq))
n_companies_south = rep(0, length(count(South_sub$Company)$freq))

for(i in 1:length(count(North_sub$Company)$freq)){
  if(count(North_sub$Company)$freq[i] == 1){
    n_companies_north[i] = count(North_sub$Company)$freq[i] 
  } else{
    n_companies_north[i] = 1
  }
}

for(i in 1:length(count(South_sub$Company)$freq)){
  if(count(South_sub$Company)$freq[i] == 1){
    n_companies_south[i] = count(South_sub$Company)$freq[i] 
  } else{
    n_companies_south[i] = 1
  }
}

sum(n_companies_north)
sum(n_companies_south)
```

Average minimum experience:

```{r}
#Use NA.omit to omit to take out NA values then use mean function to find average
count(jdata$Minimum_exp_years)
mean(North_sub$Minimum_exp_years)
mean(South_sub$Minimum_exp_years)
```

Number of Bachelor's, Master's, Ph.D:

```{r}
#Nearly same as counting number of companies:
#Create 3 vectors: n_bach, n_master, n_phd to store the values.
#For this one we need to use stringr.
#Ex: All levels -> n _bach = n_bach + 1, n_master = n_master + 1, n_phd = n_phd + 1
#BS/MS ->  n_bach = n_bach + 1, n_master = n_master + 1 (Use  stringr so that we can get only BS or only MS)
count(jdata$Standing)
Standing_sub = jdata[which(jdata$Standing != "NA"),]
North_sub_st = Standing_sub[which(Standing_sub$Region == "North CA"),]
South_sub_st = Standing_sub[which(Standing_sub$Region == "South CA"),]
Standing_sub
```

Average salary:

```{r}
#Use NA.omit to omit to take out NA values then use mean function to find average
count(jdata$Estimate_minimum_average_salary)
```

Create dataframe to plot out our histogram:

```{r}

```

Plot a histogram to see the prefer language for both prefer language 1 and prefer language 2

Plot a histogram comparing 
 
Write function to find out suitable companies: "find_company" that has 4 agrument. Minimum standing, experience, region, best language. It will return the dataframe that has all the company that fit our agruments

Write a function to predict the salary base on City, Position, Standing and Minimum experience. Simulate it around 1000 times to make sure our function correct

