---
title: "Data Related jobs"
author: "Khuong Le & Shiyong Liu"
date: July 23, 2018
---

Introduction: In this project, we performed data analysis using R language on job demand across data related field to help people better understand the basic requirement and information on data - related jobs in California State. 

We used excel to create the data set that will later be imported in r markdown file for the analysis. Moreover, we targetted job demanding social networkoing sites such as glassdoor, Indeed and Linkedin to help creating the variables for our data set. 

During the analysis, the graphs are created for visualization and  functions are created for searching. The github is used to help creating a collaborative repository to store the work files that are created in R markdown.


Load different packages that will use in this model:

```{r}
pkg_list = c("ggplot2", "plyr", "stringr")
mia_pkgs = pkg_list[!(pkg_list %in% installed.packages()[,"Package"])]
if(length(mia_pkgs) > 0) install.packages(mia_pkgs)
loaded_pkgs = lapply(pkg_list, require, character.only=TRUE)
```

Set up the data:

```{r}
jdata = read.csv("datarelatedjobs.csv")
jdata$X = NULL
names(jdata)
summary(jdata)
str(jdata)
```

Transform the data.

```{r}
jdata$Preferred_mj1 = as.character(jdata$Preferred_mj1)
jdata$Preferred_mj2 = as.character(jdata$Preferred_mj2)

for(i in 1:length(jdata$Standing)){
  if(is.na(jdata$Standing[i])){
    jdata$Standing[i] = "All levels"
  }
} 
for(i in 1:length(jdata$Preferred_mj1)){
  if(is.na(jdata$Preferred_mj1[i])){
    jdata$Preferred_mj1[i] = "Data_related"
  }  
  if(is.na(jdata$Preferred_mj2[i])){
      jdata$Preferred_mj2[i] = "Data_related"
  }
} 
count(jdata$Preferred_mj2)[order(count(jdata$Preferred_mj2)$freq, decreasing = T),]
head(jdata,10)
```

1. Comparing between companies in the North and the South of California (using side-by-side histogram):

- Number of Cities.
- Number of companies in North vs South
- Average minimum experience. 
- Average salary in North vs South

Set up number of cities.

```{r}
ct_City = count(jdata$City)
topCity = ct_City[order(ct_City$freq, decreasing = T),]

tCitydf = data.frame(City_Name =  c("S.Francisco" , "L.Angeles" , "S.Jose" , "M.View" , "S.Diego", "S.Mateo" , "P.Alto", "S.Monica", "Sunnyvale" , "RedwoodCity"),
                     NumberofPosition = c(topCity$freq[1:10]))

city_plotg = ggplot(data = tCitydf, aes(x = City_Name, y = NumberofPosition)) + 
    geom_bar(stat = "identity", position=position_dodge()) +
    geom_text(aes(label = NumberofPosition), position = position_dodge(0.95), vjust= 1.5, color = "white", size = 3)
jpeg('city_plotg.jpg')
plot(city_plotg)
city_plotg
```

Set up 2 subsets for North and South:

```{r}
North_sub = jdata[which(jdata$Region == "North CA"),]
head(North_sub,5)
South_sub = jdata[which(jdata$Region == "South CA"),]
head(South_sub,5)
```

Number of companies:

```{r}
n_companies_north = rep(0, length(count(North_sub$Company)$freq))
n_companies_south = rep(0, length(count(South_sub$Company)$freq))

for(i in 1:length(count(North_sub$Company)$freq)){
  if(count(North_sub$Company)$freq[i] == 1){
    n_companies_north[i] = count(North_sub$Company)$freq[i] 
  } else{
    n_companies_north[i] = 1
  }
}

for(i in 1:length(count(South_sub$Company)$freq)){
  if(count(South_sub$Company)$freq[i] == 1){
    n_companies_south[i] = count(South_sub$Company)$freq[i] 
  } else{
    n_companies_south[i] = 1
  }
}
sum(n_companies_north)
sum(n_companies_south)
```

Average minimum experience:

```{r}
avg_exp_N = round(mean(North_sub$Min_exp)*12,0)
avg_exp_S = round(mean(South_sub$Min_exp)*12,0)
avg_exp_N
avg_exp_S
```
 
Average salary:

```{r}
#Use NA.omit to omit to take out NA values then use mean function to find average
Omit_yr_salary_N = na.omit(North_sub$Est_min_avg_salary)
Omit_yr_salary_S = na.omit(South_sub$Est_min_avg_salary)
North_Est_min_salary = round(mean(Omit_yr_salary_N)/1000,2)
South_Est_min_salary = round(mean(Omit_yr_salary_S)/1000,2)

North_Est_min_salary
South_Est_min_salary
```

Setting up the dataframe for histogram:

```{r}
firstdf = data.frame(
          Region = rep(c("North", "South"), each = 3),
          Factor = rep(c("Average experience (in m)", "Average salary (in k)","Number Companies")),
          Number = c(avg_exp_N, North_Est_min_salary, sum(n_companies_north),
                     avg_exp_S, South_Est_min_salary, sum(n_companies_south)))

plotfirst = ggplot(data = firstdf, aes(x = Factor, y = Number, fill = Region)) + 
    geom_bar(stat = "identity", position=position_dodge()) +  
    geom_text(aes(label = Number), position = position_dodge(0.95), vjust=0.5, color = "darkred", size = 3)

jpeg('plotfirst.jpg')
plot(plotfirst)
plotfirst

```

2. Comparing between companies in the North and the South of California (using side-by-side histogram):
- Number of Bachelor's, Master's, Ph.D that companies recruiting in North vs South. (All levels = add 1 to every value)
- Top 5 most used languages.
- Top 5 major needs.

```{r}
#Nearly same as counting number of companies:
#Create 3 vectors: n_bach, n_master, n_phd to store the values.
#For this one we need to use stringr.
#Ex: All levels -> n_bach = n_bach + 1, n_master = n_master + 1, n_phd = n_phd + 1
#BS/MS ->  n_bach = n_bach + 1, n_master = n_master + 1 (Use  stringr so that we can get only BS or only MS)
Standing_sub = jdata[which(jdata$Standing != "NA"),]
North_sub_st = Standing_sub[which(Standing_sub$Region == "North CA"),]
South_sub_st = Standing_sub[which(Standing_sub$Region == "South CA"),]
```

String R, numbers of standing South:

```{r}
n_bach_S = length(str_subset(South_sub_st$Standing, pattern = "^B")) + count(South_sub_st$Standing)$freq[which(count(South_sub_st$Standing) == "All levels")]  
n_bach_S

n_master_S = length(str_subset(South_sub_st$Standing, pattern = "^MS")) + length(str_subset(South_sub_st$Standing, pattern = "MS$")) + count(South_sub_st$Standing)$freq[which(count(South_sub_st$Standing) == "All levels")] - length(str_subset(South_sub_st$Standing, pattern = "^MS$"))
n_master_S

n_PH.D_S = length(str_subset(South_sub_st$standing, pattern = "^PH.D")) + length(str_subset(South_sub_st$Standing, pattern = "PH.D$")) + count(South_sub_st$Standing)$freq[which(count(South_sub_st$Standing) == "All levels")]
n_PH.D_S
```

String R, numbers of standing North:

```{r}
n_bach_N = length(str_subset(North_sub_st$Standing, pattern = "^B")) + count(North_sub_st$Standing)$freq[which(count(North_sub_st$Standing) == "All levels")]  
n_bach_N

n_master_N = length(str_subset(North_sub_st$Standing, pattern = "^MS")) + length(str_subset(North_sub_st$Standing, pattern = "MS$")) + count(North_sub_st$Standing)$freq[which(count(North_sub_st$Standing) == "All levels")] - length(str_subset(North_sub_st$Standing, pattern = "^MS$"))
n_master_N

n_PH.D_N = length(str_subset(North_sub_st$standing, pattern = "^PH.D")) + length(str_subset(North_sub_st$Standing, pattern = "PH.D$")) + count(North_sub_st$Standing)$freq[which(count(North_sub_st$Standing) == "All levels")]
n_PH.D_N
```

Create dataframe to plot out our histogram:

```{r}
secondtdf = data.frame(
          Region = rep(c("North", "South"), each = 3),
          Factor = rep(c("Bachelor degree", "Master degree","PH.D degree")),
          Number = c(n_bach_N, n_master_N, n_PH.D_N,
                     n_bach_S, n_master_S, n_PH.D_S))

plotsecond = ggplot(data = secondtdf, aes(x = Factor, y = Number, fill = Region)) + 
    geom_bar(stat = "identity", position=position_dodge()) +  
    geom_text(aes(label = Number), position = position_dodge(0.95), vjust=0.5, color = "darkred", size = 3)

jpeg('plotsecond.jpg')
plot(plotsecond)
plotsecond
```


This graph shows the frequency of each factors in the standing variable of jdata set. Each factor is subsetted by region, which is north vs south in California state. The indepedent variables are bachelor degree, master degree and Ph.D.The numbers on top of each bar indicates the frequency of that factor. The color indicates region difference.

Write function to count and put them in decreasing order:

Intputs: 2 columns of language variables that needs to be counted ( North vs South) and Numbers of different language.
Output: Dataframe in order of your choice (decreasing/ increasing)

```{r}
count_order = function(col1, col2, order, top){
  ct1 = count(col1)
  ct2 = count(col2)
  if(order == "decreasing"){
    ct1_inc = ct1[order(ct1$freq, decreasing = T),]
    ct2_inc = ct2[order(ct2$freq, decreasing = T),]
    
    cl = length(ct1_inc$freq) - length(ct2_inc$freq)

    if(cl < 0){
      fresult = data.frame(Lang1 = c(as.character(ct1_inc$x), rep(NA,abs(cl))),
                            Freq1 = c(ct1_inc$freq, rep(NA,abs(cl))),
                            Lang2 = as.character(ct2_inc$x),
                            Freq2 = ct2_inc$freq)                  

    } else{
      fresult = data.frame(Lang1 = as.character(ct1_inc$x),
                            Freq1 = ct1_inc$freq,
                            Lang2 = c(as.character(ct2_inc$x), rep(NA,cl)),
                            Freq2 = c(ct2_inc$freq,rep(NA,cl)))         
      }
  }
  else{
    ct1_dec = ct1[order(ct1$freq, decreasing = F),]
    ct2_dec = ct2[order(ct2$freq, decreasing = F),]
    
    cl = length(ct1_dec$freq) - length(ct2_dec$freq)

    if(cl < 0){
      fresult = data.frame(Lang1 = c(as.character(ct1_dec$x), rep(NA,abs(cl))),
                            Freq1 = c(ct1_dec$freq, rep(NA,abs(cl))),
                            Lang2 = as.character(ct2_dec$x),
                            Freq2 = ct2_dec$freq)                  
    } else{
      fresult = data.frame(Lang1 = as.character(ct1_dec$x),
                            Freq1 = ct1_dec$freq,
                            Lang2 = c(as.character(ct2_dec$x), rep(NA,cl)),
                            Freq2 = c(ct2_dec$freq,rep(NA,cl))) 
    }
  }
  for(i in 1:15){
    for(j in 1:15){
      if(fresult$Lang1[i] == fresult$Lang2[j]){
        fresult$Freq1[i] = fresult$Freq1[i] + fresult$Freq2[j]
      }
    }
  }
  if(order == "decreasing"){
    res = fresult[order(fresult$Freq1, decreasing = T),]
  } else {
    res = fresult[order(fresult$Freq1, decreasing = F),]
  }
  return(res[1:top,1:2])
}

```


Plot a histogram to see the prefer language for both prefer language 1 and prefer language 2
Dataframe for counting Language:

```{r}
sl = count_order(South_sub_st$Preferred_lang1, South_sub_st$Preferred_lang2,"decreasing", 5)
nl = count_order(North_sub_st$Preferred_lang1, North_sub_st$Preferred_lang2,"decreasing", 5)
sl
nl
thirddf = data.frame(
          Region = rep(c("North", "South"), each = 5),
          Factor = rep(c("SQL", "Python", "R", "Java", "Tableau - C/C++"),2),
          Number = c(nl$Freq1[1], nl$Freq1[2], nl$Freq1[3], nl$Freq1[4], nl$Freq1[5], 
                     sl$Freq1[1], sl$Freq1[2], sl$Freq1[3], sl$Freq1[4], sl$Freq1[5]))

plotthird = ggplot(data = thirddf, aes(x = Factor, y = Number, fill = Region)) + 
    geom_bar(stat = "identity", position=position_dodge()) +  
    geom_text(aes(label = Number), position = position_dodge(0.95), vjust=0.5, color = "darkred", size = 3)

jpeg('plotthird.jpg')
plot(plotthird)
plotthird
```

The graph shows the frequency of top 5 most demanding language in north vs south California state. By calculating each language frequency, we conclude that Java, Python,R,SOL and Tableau are top 5 most demanding language in North California State, where as Java, Python,R,SOL and c/c++ are most demanding language in South California State. Each language is subsetted by the north and south region in California State. The numbers on top of each bar indicate the frequency of that Language.

Calculate the frequency of different Language in different way to double check:

```{r}
n_cpp_North = length(str_subset(North_sub_st$Preferred_lang1, pattern = "^C\\++$")) +length(str_subset(North_sub_st$Preferred_lang2, pattern = "C\\++$"))

n_cpp_South = length(str_subset(South_sub_st$Preferred_lang1, pattern = "^C\\++$")) +length(str_subset(South_sub_st$Preferred_lang2, pattern = "C\\++$"))

n_java_South = length(str_subset(South_sub_st$Preferred_lang1, pattern = "^Java$")) +length(str_subset(South_sub_st$Preferred_lang2, pattern = "^Java$"))

n_java_North = length(str_subset(North_sub_st$Preferred_lang1, pattern = "^Java$")) +length(str_subset(North_sub_st$Preferred_lang2, pattern = "^Java$"))

n_Python_North = length(str_subset(North_sub_st$Preferred_lang1, pattern = "^Python$"))+length(str_subset(North_sub_st$Preferred_lang2, pattern = "^Python$"))

n_Python_South = length(str_subset(South_sub_st$Preferred_lang1, pattern = "^Python$"))  +length(str_subset(South_sub_st$Preferred_lang2, pattern ="^Python$"))

n_R_North = length(str_subset(North_sub_st$Preferred_lang1, pattern = "^R$")) +length(str_subset(North_sub_st$Preferred_lang2, pattern = "R$"))

n_R_South = length(str_subset(South_sub_st$Preferred_lang1, pattern = "^R$"))  +length(str_subset(South_sub_st$Preferred_lang2, pattern = "R$"))
n_R_South
n_SQL_North = length(str_subset(North_sub_st$Preferred_lang1, pattern = "^SQL$")) +length(str_subset(North_sub_st$Preferred_lang1, pattern = "/SQL$")) +length(str_subset(North_sub_st$Preferred_lang2, pattern = "^SQL$"))

n_SQL_South = length(str_subset(South_sub_st$Preferred_lang1, pattern = "^SQL$"))+ length(str_subset(South_sub_st$Preferred_lang1, pattern = "/SQL$"))+ length(str_subset(South_sub_st$Preferred_lang2, pattern = "^SQL$"))


res = cbind(nl,sl)
handwaydf = data.frame(NLang1 = c("SQL", "Python", "R", "Java", "C/C++"),
                       NFreg1 = c(n_SQL_North, n_Python_North, n_R_North, n_java_North, n_cpp_North),
                       SLang1 = c("SQL", "Python", "R", "Java", "C/C++"),
                       NLang1 = c(n_SQL_South, n_Python_South, n_R_South, n_java_South, n_cpp_South))
res
handwaydf
```

Plot a histogram to see the prefer major for both prefer major 1 and prefer major 2
Dataframe for counting Major:


```{r}
smj = count_order(South_sub_st$Preferred_mj1, South_sub_st$Preferred_mj2,"decreasing", 5)
nmj = count_order(North_sub_st$Preferred_mj1, North_sub_st$Preferred_mj2,"decreasing", 5)
smj
nmj
forthdf = data.frame(
          Region = rep(c("North", "South"), each = 5),
          Factor = rep(c("CS", "Tech field - Stats", "Stats - Tech field", "Math", "Engineering"),2), 
          Number = c(nmj$Freq1[1], nmj$Freq1[2], nmj$Freq1[3], nmj$Freq1[4], nmj$Freq1[5], 
                     smj$Freq1[1], smj$Freq1[2], smj$Freq1[3], smj$Freq1[4], smj$Freq1[5]))

plotforth = ggplot(data = forthdf, aes(x = Factor, y = Number, fill = Region)) + 
    geom_bar(stat = "identity", position=position_dodge()) +  
    geom_text(aes(label = Number), position = position_dodge(0.95), vjust=0.5, color = "darkred", size = 3)

jpeg('plotforth.jpg')
plot(plotforth)
plotforth


```

Combining 1 and 2 together:

```{r}
combinedf = data.frame(
          Region = rep(c("North", "South"), each = 6),
          Factor = c("XP_N(m)", "SalN(k)", "#CompN", "#Ba_N", "#Ma_N", "#PHD_N", "XP_S(m)", "SalS(k)", "#CompS", "#Ba_S", "#Ma_S", "#PHD_S"),
          Number = c(avg_exp_N, North_Est_min_salary, sum(n_companies_north), n_bach_N, n_master_N, n_PH.D_N, avg_exp_S, South_Est_min_salary, sum(n_companies_south), n_bach_S, n_master_S, n_PH.D_S))


plotcombine = ggplot(data = combinedf, aes(x = Factor, y = Number, fill = Region)) + 
    geom_bar(stat = "identity", position=position_dodge()) +  
    geom_text(aes(label = Number), position = position_dodge(0.95), vjust=0.5, color = "darkred", size = 3)
jpeg('plotcombine.jpg')
plot(plotcombine)
plotcombine
```

3 - Write function to find out suitable companies: "find_company" that has 4 agrument. Minimum standing, experience, region, best language. It will return the dataframe that has all the company that fit our agruments

Let's just use the top 20 preferred language:

```{r}
top20 = count_order(jdata$Preferred_lang1, jdata$Preferred_lang2,"decreasing", 20)
top20
```

Change the data in the way that if "Bachelor"/"All levels" is able to be looked at means Easy, "Master" means Medium and "PHD" will be Hard.

Our function will now base on Difficulty, Experience, Region, Best Language:


```{r}
for (i in 1:length(jdata$Standing)){
  if(!is.na(str_extract(jdata$Standing[i], pattern = "^B")) || !is.na(str_extract(jdata$Standing[i], pattern = "^A"))){
    jdata$Difficulty[i] = "Easy"
  } else if (!is.na(str_extract(jdata$Standing[i], pattern = "^M"))){
    jdata$Difficulty[i] = "Medium"
  } else if (!is.na(str_extract(jdata$Standing[i], pattern = "^P"))){
    jdata$Difficulty[i] = "Hard"
  }
}
```

Function find_company using intersect and union subsets. The input of the function will be Difficulty, Minimum years of experience, region and language in Jdata set. The output will be the list of companies name that satisfied the arguments.

```{r}
find_company = function(Difficulty, Exp, Region,Language){
  Argu_y = jdata$Company[which(jdata$Difficulty == Difficulty)]
  Argu_x = jdata$Company[which(jdata$Min_exp <= Exp)]
  Argu_z = jdata$Company[which(jdata$Region == Region)]
  Argu_u1 = jdata$Company[which(jdata$Preferred_lang1 == Language)]
  Argu_u2 = jdata$Company[which(jdata$Preferred_lang2 == Language)]
  
  return(union(intersect(intersect(Argu_x,Argu_y),intersect(Argu_z,Argu_u1)),intersect(intersect(Argu_x,Argu_y),intersect(Argu_z,Argu_u2))))

}

find_company("Medium", 5, "South CA","R")
```

Another way using subset find_company:

```{r}
n_find_company = function(Difficulty, Exp, Region,Language){
a = jdata$Company[which(jdata$Difficulty == Difficulty),]
b = jdata$Company[which(jdata$Min_exp <= Exp),]
c = jdata$Company[which(jdata$Region == Region),]
d1 = jdata$Company[which(jdata$Preferred_lang1 == Language),]
d2 = jdata$Company[which(jdata$Preferred_lang2 == Language),]

pre_r1 = a[which(a$Min_exp <= Exp)]
nex_r1 = pre_r1[which(pre_r1$Region == Region)]
r1     = nex_r1[which(nex_r1$Preferred_lang1 == Language)]

pre_r2 = a[which(a$Min_exp <= Exp)]
nex_r2 = pre_r2[which(pre_r2$Region == Region)]
r2     = nex_r2[which(nex_r2$Preferred_lang2 == Language)]

rbind(r1,r2)
}
find_company("Medium", 5, "South CA","R")
```

4- Comparing between Data Scientist and Data Engineer.
Base on jdata.
- Number of Data Scientist vs number of Data Engineer 
- Top 3 language for Data Scientist vs Data Engineer 
- Best est_min_avg salary for each
- City that offer the most Data Scientist position vs Data Engineer.

Finding total number of Data Engineers and Data Scientists in North and South of California state using String R:

```{r}
major = count(jdata$Position)
count(North_sub_st$Position)

n_eng_north = length(str_subset(North_sub_st$Position, pattern = "^Data Engineer"))+length(str_subset(North_sub_st$Position, pattern = "Data Engineer$")) - length(str_subset(North_sub_st$Position, pattern = "^Data Engineer$"))

n_eng_north

n_eng_south = length(str_subset(South_sub_st$Position, pattern = "^Data Engineer"))+length(str_subset(South_sub_st$Position, pattern = "Data Engineer$")) - length(str_subset(South_sub_st$Position, pattern = "^Data Engineer$"))

n_eng_south

n_sci_north =  length(str_subset(North_sub_st$Position, pattern = "^Data Scientist"))+length(str_subset(North_sub_st$Position, pattern = "Data Scientist$")) - length(str_subset(North_sub_st$Position, pattern = "^Data Scientist$"))

n_sci_north

n_sci_south = length(str_subset(South_sub_st$Position, pattern = "^Data Scientist"))+length(str_subset(South_sub_st$Position, pattern = "Data Scientist$")) - length(str_subset(South_sub_st$Position, pattern = "^Data Scientist$"))

n_sci_south
```


Comparing between numbers of data engineers and data scientists in the North and the South of California (using side-by-side histogram):

```{r}
P_data_fr = data.frame(Region = c("North","South","North","South"),
  Number = c(n_eng_north,n_eng_south,n_sci_north,n_sci_south),
  Factor = rep(c("Data Engineer","Data Scientist"),each = 2)
)

Plot_1 = ggplot(data = P_data_fr, aes(x = Factor, y = Number, fill = Region)) + 
    geom_bar(stat = "identity", position=position_dodge()) +  
    geom_text(aes(label = Number), position = position_dodge(0.95), vjust=0.5, color = "darkred", size = 3)

#jpeg('Plot_1.jpg')
#plot(Plot_1)
Plot_1
```

This histogram shows the comparison between the number of data engineer and data scientist in north and south of California. The independent variables are data engineer and data scientist, and the number on top of them represent their frequency.

Top 3 language for Data Scientist vs Data Engineer 

Set up 2 subsets for Data Scientist and Engineer in California State:

```{r}
Standing_sub = jdata[which(jdata$Standing != "NA"),]
Eng_sub = Standing_sub[which(Standing_sub$Position == "Data Engineer"),]
Sci_sub = Standing_sub[which(Standing_sub$Position == "Data Scientist"),]
```

Number of top 3 most used lanuage (Python, R, SQL) in Data Engineer vs Data Scientist position:

```{r}
n_Python_Eng = length(str_subset(Eng_sub$Preferred_lang1, pattern = "Python"))+length(str_subset(Eng_sub$Preferred_lang2, pattern = "Python"))

n_Python_Eng

n_Python_Sci = length(str_subset(Sci_sub$Preferred_lang1, pattern = "Python"))+length(str_subset(Sci_sub$Preferred_lang2, pattern = "Python"))

n_Python_Sci

n_R_Eng = length(str_subset(Eng_sub$Preferred_lang1, pattern = "R"))+length(str_subset(Eng_sub$Preferred_lang2, pattern = "R"))

n_R_Eng

n_R_Sci = length(str_subset(Sci_sub$Preferred_lang1, pattern = "R"))+length(str_subset(Sci_sub$Preferred_lang2, pattern = "R"))

n_R_Sci

n_SQL_Eng = length(str_subset(Eng_sub$Preferred_lang1, pattern = "SQL"))+length(str_subset(Eng_sub$Preferred_lang2, pattern = "SQL"))

n_SQL_Eng

n_SQL_Sci = length(str_subset(Sci_sub$Preferred_lang1, pattern = "SQL"))+length(str_subset(Sci_sub$Preferred_lang2, pattern = "SQL"))

n_SQL_Sci
```

Comparing between number of Python, R, SQL in the position of Data Scientist and Data Engineer in California (using side-by-side histogram):                 

```{r}
Lang_data_fr = data.frame( Factor =c("Python","R","SQL"),
      Position = rep(c("Data Scientist", "Data Engineer"),each = 3),
      Number =  c(n_Python_Sci,n_R_Sci,n_SQL_Sci,n_Python_Eng,n_R_Eng,n_SQL_Eng))

Plot_position = ggplot(data = Lang_data_fr, aes(x = Factor, y = Number, fill = Position)) + 
    geom_bar(stat = "identity", position=position_dodge()) +  
    geom_text(aes(label = Number), position = position_dodge(0.95), vjust=0.5, color = "darkred", size = 3)

#jpeg('Plot_position.jpg')
#plot(Plot_position)
Plot_position

```

This histogram shows the number of top 3 most used language in the position of data scientist and data engineer. The histogram is subsetted by postion of data scientist and data engineer, and the number on top of each bars represents the frequency of different indepedent variables.  

Best estimated minimum salary for Data scientist and Data Engineer

Average Salary
```{r}
#Use NA.omit to omit to take out NA values then use mean function to find average Salary 

Omit_yr_salary_N = na.omit(Eng_sub$Est_min_avg_salary)
Omit_yr_salary_S = na.omit(Sci_sub$Est_min_avg_salary)
Eng_Est_min_salary = round(mean(Omit_yr_salary_N)/1000,2)
Sci_Est_min_salary = round(mean(Omit_yr_salary_S)/1000,2)

Eng_Est_min_salary
Sci_Est_min_salary
```

Comparing between estimated minimum yearly salary (in thousands) for Data scientist and Data Engineer in California (using side-by-side histogram): 

```{r}
sala_data_frame = data.frame(
  Position = rep(c("Data Scientist", "Data Engineer"),each = 1),
  Number = c(Sci_Est_min_salary,Eng_Est_min_salary),
  Factor = c("Data Scientist Salary","Data Engineer Salary"))
            
Plot_Salary = ggplot(data = sala_data_frame, aes(x = Factor, y = Number, fill = Position)) + 
    geom_bar(stat = "identity", position=position_dodge()) +  
    geom_text(aes(label = Number), position = position_dodge(0.95), vjust=0.5, color = "darkred", size = 3)

#jpeg('Plot_Salary.jpg')
#plot(Plot_Salary)
Plot_Salary
                             
```

This histogram shows the comparison between estimated minimum yearly salary (in thousands) for Data scientist and Data Engineer in California. We subsetted the histogram by postion of data scientist and data engineer in the data set.The number on top of each bars represents the frequency of different independent variables. 

Finding the top 3 cities that have both most number of data Engineers and data Scientists:

```{r}
intersect(Sci_sub$City,Eng_sub$City)
count(Sci_sub$City)
count(Eng_sub$City)

City_Eng_Sci = c("San Francisco","Santa Monica","Los Angeles")
```

Number of Data Engineer and Data Scientist in San Francisco, Santa Monica and Los Angeles of California:

```{r}
n_Eng_San = length(str_subset(Eng_sub$City, pattern = "San Francisco"))

n_Eng_San

n_Sci_San = length(str_subset(Sci_sub$City, pattern = "San Francisco"))

n_Sci_San

n_Eng_Mon = length(str_subset(Eng_sub$City, pattern = "Santa Monica"))

n_Eng_Mon

n_Sci_Mon = length(str_subset(Sci_sub$City, pattern = "Santa Monica"))

n_Sci_Mon

n_Eng_Los = length(str_subset(Eng_sub$City, pattern = "Los Angeles"))

n_Eng_Los

n_Sci_Los = length(str_subset(Sci_sub$City, pattern = "Los Angeles"))

n_Sci_Los

```

Comparing number of Data Engineer and Data Scientist in San Francisco, Santa Monica and Los Angeles of California:(using side-by-side histogram): 

```{r}
City_Data_fr = data.frame (City = rep(c("San Francisco","Santa Monica","Los Angeles"),each = 2),
                    Factor = c("Data Scientist","Data Engineer"),
                    Number = c(n_Sci_San,n_Eng_San,n_Sci_Mon,n_Eng_Mon,n_Sci_Los,n_Eng_Los))

Plot_3 = ggplot(data = City_Data_fr, aes(x = Factor, y = Number, fill = City)) + 
    geom_bar(stat = "identity", position=position_dodge()) +  
    geom_text(aes(label = Number), position = position_dodge(0.95), vjust=0.5, color = "darkred", size = 3)

#jpeg('Plot_3.jpg')
#plot(Plot_3)
Plot_3
                
```

This histogram shows the comparison of number of Data Engineer and Data Scientist in San Francisco, Santa Monica and Los Angeles of California. In here, we subsetted the histogram by City of San Francisco,Santa Monica and Los Angeles and
made data engineer and data scientist as two of the independent variables.The number on top of each bars represents the frequency of different indepedent variables. 

5 - Linear regression model to find our salary and check if it fit.

Data transformation:
Minimum experience

```{r}
for(i in 1:length(jdata$Min_exp)){
  if(jdata$Min_exp[i] == 1 & !is.na(jdata$Est_min_avg_salary[i])){
    if(jdata$Est_min_avg_salary[i] >= 150000){
      jdata$Min_exp[i] = 4
    }
    else if(jdata$Est_min_avg_salary[i] >= 130000 &
            jdata$Est_min_avg_salary[i] <  150000){
      jdata$Min_exp[i] = 3
    }
    else if (jdata$Est_min_avg_salary[i] >= 100000 &
             jdata$Est_min_avg_salary[i] <  130000){
      jdata$Min_exp[i] = 2
    }
    else{jdata$Min_exp[i] = 1}
  } 
}
```

Add variable names: Jobs density to show which city offers lots of position, which city offers less position.

```{r}
set.seed(2018)
result = jdata$Est_min_avg_salary[which.min(jdata$Est_min_avg_salary)]
for (i in 1:length(jdata$City))
if(count(jdata$City)$freq[which(count(jdata$City)$x == (jdata$City[i]))] > mean(count(jdata$City)$freq)){ 
  jdata$Jobs_density[i] = "High"
  jdata$City_Bonus[i] = round(rnorm(1, mean = result * 1.5, sd = 1500),1) 
} else {
  jdata$Jobs_density[i] = "Low"
  jdata$City_Bonus[i] = round(rnorm(1, mean = result * 1.45, sd = 1500),1) 
}
```

Set up bonuses for salary:

According to job function, we can divide position into 3 functions (Engineer, Scientist, Analyst). Assume bonuses follow a normal distribution that has mean = 7000 - 6000 - 5000 and sd = 500

```{r}
set.seed(2018)
for (i in 1:length(jdata$Position)){
  if(str_detect(jdata$Position[i], pattern = "Engineer")){
    jdata$Posi_Bonus[i] = round(rnorm(1, mean = 25000, sd = 1000),1)
  }  else if (str_detect(jdata$Position[i], pattern = "Scientist")){
    jdata$Posi_Bonus[i] = round(rnorm(1, mean = 22000, sd = 1000),1)
  }  else if (str_detect(jdata$Position[i], pattern = "Analyst")){
    jdata$Posi_Bonus[i] = round(rnorm(1, mean = 15000, sd = 1000),1)
  } else {
    jdata$Posi_Bonus[i] = round(rnorm(1, mean = 10000, sd = 1000),1)
  }
}

```

Finding bonuses for with the year of experience:

```{r}
calc_salary = na.omit(jdata$Est_min_avg_salary)
avg_salary = mean(calc_salary[which(calc_salary >= 100000)])
salaryvsyear = lm(Est_min_avg_salary ~ Min_exp, data = jdata)
summary(salaryvsyear)
#Since the p-value for this model is very small it looks like the Min_exp is significant.
exp_multiplier = summary(salaryvsyear)$coef[2,1]
```

Set up bonus for Bonuses with years experience

```{r}
set.seed(2018)
for (i in 1:length(jdata$Min_exp)){
  if(is.na(jdata$Est_min_avg_salary[i])){
    jdata$Exp_Bonus[i] = round(exp_multiplier * jdata$Min_exp[i] / mean(jdata$Min_exp) ,1) 
  }  else if(!is.na(jdata$Est_min_avg_salary[i]) & jdata$Est_min_avg_salary[i] >= 100000){
    jdata$Exp_Bonus[i] = round(exp_multiplier * jdata$Min_exp[i] ,1) 
  }  else {
    jdata$Exp_Bonus[i] = round(exp_multiplier * jdata$Min_exp[i] /2 ,1) 
  }
}
head(jdata,50)

```

Original model:

Compare between 
```{r}
par(mfrow = c(1,3))
model1 = lm(Est_min_avg_salary ~ City_Bonus, data = jdata)
plot(jdata$Est_min_avg_salary ~ jdata$City_Bonus, main = "Salary vs City Bonus", xlab = "City Bonus", ylab = "Salary")
abline(model1)

model2 = lm(Est_min_avg_salary ~ Posi_Bonus, data = jdata)
plot(jdata$Est_min_avg_salary ~ jdata$Posi_Bonus, main = "Salary vs Position Bonus", xlab = "Position Bonus", ylab = "Salary")
abline(model2)

model3 = lm(Est_min_avg_salary ~ Exp_Bonus, data = jdata)
plot(jdata$Est_min_avg_salary ~ jdata$Exp_Bonus, main = "Salary vs Experience Bonus", xlab = "Experience Bonus", ylab = "Salary")
abline(model3)
```

New model:

```{r, warning = FALSE}
fit_model = lm(Est_min_avg_salary ~ Exp_Bonus + City_Bonus + Posi_Bonus, data = jdata)
summary(fit_model)
```

For the model of Est_min_avg_salary and the bonuses, beside City_Bonus is 0.2611, the order value is very small
Using backward aic to find out the selected model:

```{r}
full_model = lm(Est_min_avg_salary ~ Min_exp + Exp_Bonus + City_Bonus + Posi_Bonus, data = jdata)
sele_model = step(full_model, direction = "backward", trace = 0)
sele_model

```

From the select model, it looks like City Bonus is not really great for the model.

```{r}
for(i in 1:length(jdata$Est_min_avg_salary)){
  jdata$New_min_avg_salary[i] = jdata$City_Bonus[i] + jdata$Posi_Bonus[i] + jdata$Exp_Bonus[i]
}

```

Compare new salary and estimate one.

```{r}
model_compare = lm(Est_min_avg_salary ~ New_min_avg_salary, data = jdata)
plot(jdata$Est_min_avg_salary, jdata$New_min_avg_salary)
abline(model_compare)
```

It looks like the from the model, the abline is not following what the data has. 

Logistic Regression model:

Set up logistic model. 
```{r}
for(i in 1:length(jdata$Region)){
  if(jdata$Region[i] == "North CA"){
    jdata$Region_Logic[i] = 1
  } else { 
    jdata$Region_Logic[i] = 0
  }
}
head(jdata,50)

```

The model:
```{r, warning = FALSE}
fit_glm = glm(Region_Logic ~ Est_min_avg_salary, data = jdata, family = binomial)
fit_lm  =  lm(Region_Logic ~ Est_min_avg_salary, data = jdata)

qplot(jdata$Est_min_avg_salary, jdata$Region_Logic) + stat_smooth()
```

Find interactions between Region_Logic and the bonuses.

```{r}
reg_interaction = glm(Region_Logic ~ Est_min_avg_salary + City_Bonus + Posi_Bonus + Exp_Bonus + Est_min_avg_salary:City_Bonus, family = binomial, data = jdata)
summary(reg_interaction) 
```


Conclusion:

By completing this project, we are able to understand more in depth about the basic requirements and information about data - related job in the state of California, in that we are able to create  graphs to visualize the behavior of different variables from the data set and write out functions to help obtaining the outputs that would match the conditional inputs.  



Body: Talk about all of the graphs we have.
- What's the graph about.
- Why do we make this graph
- Interesting number, reasons?
Talk about those functions we made

